

```{r}
#data
load("~/470calorimetry/x.RData")
```

```{r}
#packages
library(lme4)
library(pbkrtest)
library(xtable)
```

```{r}
rownames(x)<-as.character(1:nrow(x))
tabl1<-x[,c("ParticipantID","Measurement","REE","REEperkg","BMI","obese",
           "ventilator","ICUorSDU")]
colnames(tabl1)<-c("ID","Measurement","REE","REEperKG","BMI","Obese",
                 "Ventilator","ICUorSDU")
xtabl1<- xtable(tabl1,caption="Calorimetry data structured in `long format' for mixed model analysis.",
               digits=c(0,0,0,0,2,2,0,0,0))
label(xtabl1)<-"tabl:data"
print(xtabl1,scalebox=0.7)
```


\subsection{Location (ICU/SDU) Analysis}

\subsubsection{Resting Energy Expenditure (REE)}

In this section, we fit three new LMMs to help us find overall and ICU/SDU status specific estimates for mean REE as well as test for a difference in REE between ICU and SDU patients. The first model includes just participant ID as a random effect ((\texttt{REE$\sim$(1|ParticipantID)} as an \texttt{R}
model formula). The second model includes ICU/SDU status as a fixed effect, and participant ID as a random effect ((\texttt{REE$\sim$ICUorSDU+(1|ParticipantID)} as an \texttt{R}
model formula). The third model includes obesity and ICU/SDU status as fixed effects, and participant ID as a random effect ((\texttt{REE$\sim$obese+ICUorSDU(1|ParticipantID)} as an \texttt{R}
model formula).

```{r}
REE.icu0<-lmer(REE~(1|ParticipantID),data=x)
REE.icu1<-lmer(REE~ICUorSDU+(1|ParticipantID),data=x)
REE.icu.obesity<-lmer(REE~ICUorSDU + obese + (1|ParticipantID),data=x)
```


```{r}
Again, we assume that the error and random effects are approximately normally distributed, and we will use Q--Q plots to check for violations.
```


```{r}
re.out<-ranef(REE.icu1)
re.out<-re.out$ParticipantID[[1]]
err.out<-residuals(REE.icu1)
par(mfrow=c(1,2),las=1,cex.axis=1.0,cex.lab=1.0,
   cex.sub=1.5,pch=16,cex=0.5,cex.main=1.25)
qqnorm(re.out,main="QQ Plot of ICU Analysis Patient REs\n REE Analysis")
abline(a=0,b=sd(re.out),lwd=2,col=2)
qqnorm(err.out,main="QQ Plot of ICU Analysis Residuals\n REE Analysis")
abline(a=0,b=sd(err.out),lwd=2,col=2)

re.out<-ranef(REE.icu.obesity)
re.out<-re.out$ParticipantID[[1]]
err.out<-residuals(REE.icu.obesity)
par(mfrow=c(1,2),las=1,cex.axis=1.0,cex.lab=1.0, cex.sub=1.5,pch=16,cex=0.5,cex.main=1.25)
qqnorm(re.out,main="QQ Plot of ICU+Obese Analysis Patient REs\n REE Analysis")
abline(a=0,b=sd(re.out),lwd=2,col=2)
qqnorm(err.out,main="QQ Plot of ICU+Obese Analysis Residuals\n REE Analysis")
abline(a=0,b=sd(err.out),lwd=2,col=2)

REEicu.kr<-KRmodcomp(REE.icu0,REE.icu1)
REEicu.kr
ndi0<-expand.grid(ICUorSDU="SDU",ParticipantID="L0000")
REEicuFit0<-predict(REE.icu0,re.form=NA,se.fit=TRUE,newdata=ndi0)
ndi1<-expand.grid(ICUorSDU=c("SDU","ICU"),ParticipantID="L0000")
REEicuFit1<-predict(REE.icu1,re.form=NA,se.fit=TRUE,newdata=ndi1)
```

{\textbf{Findings: REE by ICU/SDU Status.}}  We use the LMM that includes
ICU/SDU status as a main effect and patient ID as a random effect to estimate
ICU/SDU status specific mean REE.  We estimate the mean REE among SDU
patients to be \rinline{round(REEicuFit1$fit[1],2)} (SE =
\rinline{round(REEicuFit1$se.fit[1],2)}) and among ICU patients
to be \rinline{round(REEicuFit1$fit[2],2)} (SE =
\rinline{round(REEicuFit1$se.fit[2],2)}).  We use the
Kenward--Roger adjusted F--test comparing the null model to the model with ICU/SDU status as a fixed effect to test for a difference in REE by ICU/SDU status.  This results in a p--value of
\rinline{REEicu.kr$stats$p.value}.


```{r}
ndio0<-expand.grid(ICUorSDU=c("SDU","ICU"),obese = c(0, 1), ParticipantID="L0000")
REEicuobesityFit1<-predict(REE.icu.obesity,re.form=NA,se.fit=TRUE,newdata=ndio0)
```

{\textbf{Findings: REE by ICU/SDU and Obesity Status.}}  We use the LMM that includes ICU/SDU status and obesity status as main effects and patient ID as a random effect to estimate ICU/SDU and obesity status--specific mean REE.  We estimate the mean REE among non--obese patients in the SDU to be \rinline{round(REEicuobesityFit1$fit[1],2)} (SE = \rinline{round(REEicuobesityFit1$se.fit[1],2)}). Among non--obese patients in the ICU, we estimate the mean REE to be \rinline{round(REEicuobesityFit1$fit[2],2)} (SE = \rinline{round(REEicuobesityFit1$se.fit[2],2)}). Among obese patients in the SDU, we estimate the mean REE to be \rinline{round(REEicuobesityFit1$fit[3],2)} (SE = \rinline{round(REEicuobesityFit1$se.fit[3],2)}). Finally, among obese patients in the ICU, we estimate the mean REE to be \rinline{round(REEicuobesityFit1$fit[4],2)} (SE = \rinline{round(REEicuobesityFit1$se.fit[4],2)}).

```{r}
% REE.icuobesity.kr <- KRmodcomp(REE.obesity1, REE.icu.obesity)
% REE.icuobesity.kr
```



We use the Kenward--Roger adjusted F--test comparing the previously defined LMM with only the patient random effect and the obesity main effect to the model that has a random effect for patient and main effects for obesity and ICU/SDU status (alternative model) to test for a difference in REE by ICU/SDU status when adjusting for obesity.  This results in a p--value of
\rinline{REE.icuobesity.kr$stats$p.value}.


```{r}
REE.icuobesity.kr2 <- KRmodcomp(REE.icu1, REE.icu.obesity)
REE.icuobesity.kr2
```



We use the Kenward--Roger adjusted F--test comparing the LMM with only the patient random effect and the ICU/SDU main effect to the model that has a random effect for patient and main effects for obesity and ICU/SDU status (alternative model) to test for a difference in REE by obesity status when adjusting for ICU/SDU status.  This results in a p--value of
\rinline{REE.icuobesity.kr2$stats$p.value}.


